\documentclass[11pt]{article}
\usepackage{geometry}                
\geometry{letterpaper,tmargin=1in,bmargin=1in,lmargin=1in,rmargin=1in} 
\usepackage[parfill]{parskip}    % Activate to begin paragraphs with an empty line rather than an indent
\usepackage{graphicx}
\usepackage{amsmath, amsfonts, amsthm, amssymb} 
\usepackage[shortlabels]{enumitem}
\usepackage{xcolor}

\parskip = 0.1in

\pagestyle{myheadings}
\markright{Homework of IOE 611 Nonlinear Programming\hfill Yulun Zhuang \hfill}

\input{611defs}

\newcommand{\oh}{\frac12}
\newcommand{\st}{\text{subject to}}
\newcommand{\gfb}{\nabla f(\bar x)}
\newcommand{\hfb}{H(\bar x)}

\newtheorem{theorem}{Theorem}[section]
\newtheorem{remark}[theorem]{Remark}%[section]
\newtheorem{definition}[theorem]{Definition}%[section]
\newtheorem{proposition}[theorem]{Proposition}%[section]
\newtheorem{lemma}[theorem]{Lemma}%[section]
\newtheorem{corollary}[theorem]{Corollary}%[section]
\newtheorem{assumption}{Assumption}
\newtheorem{claim}{Claim}
\newtheorem{exam}{Example}
\newenvironment{solution}
  {\renewcommand\qedsymbol{$\square$}\begin{proof}[\textbf{Solution}]}
  {\end{proof}}
\renewcommand{\proofname}{\textbf{Proof}}

\newcommand{\red}[1]{\textcolor{red}{#1}}
\newcommand{\blue}[1]{\textcolor{blue}{#1}}
\newcommand{\green}[1]{\textcolor{green}{#1}}
\newcommand{\grad}{\nabla}
\newcommand{\hess}{\nabla^2}
\newcommand{\tr}{\text{tr}}

\newcommand{\dd}{\mathrm{d}}
\newcommand{\RR}{\mathbb{R}}
\newcommand{\NN}{\mathbb{N}}
\newcommand{\ZZ}{\mathbb{Z}}
\newcommand{\bS}{\mathbb{S}}

\newcommand{\pd}[2][]{ \frac{\partial #1}{\partial #2}} % Partial derivatives
\renewcommand{\d}{{\rm d}}
\newcommand{\ddt}{\frac{\d}{\d t}}
\newcommand{\half}{\frac{1}{2}}
\newcommand{\inv}{^{-1}}
\newcommand{\T}{^\top}

\begin{document}
\title{IOE 611: Midterm}
\author{Yulun Zhuang}
\maketitle
%**********************************
\section*{Problem 1}
Obtain the dual cone of the set $K \subset \real^3$ defined as:
\[
K = \{(x, y, z):y>0, ye^{x/y} \leq z\}
\]

\begin{solution}
    TODO
\end{solution}



\clearpage
\section*{Problem 2}
Consider the following optimization problem:

\begin{align}
    \min_{x\in\real^n}\quad &f(x)\notag\\
    \text{s.t.}\quad 
    &\ones\T x = 1\tag{P1}\\
    &x\succeq 0\notag
\end{align}
where $\ones$ is a vector of ones, i.e. $\ones = [1, \dots, 1]\T\in\real^n$. Assume that $f(x)$ is convex and differentiable.

(a) Prove that $x^*$ is optimal for (P1) if and only if $x^*$ is feasible and
\[
\min_{1\leq i \leq n} \grad f(x^*)_i \geq \grad f(x^*)\T x^*
\]
where $\grad f(x^*)_i$ is the ith element of $\grad f(x^*)$.

(b) Prove that $x^*$ is optimal for (P1) if and only if $x^*$ is feasible and, for each $k$,
\[
x_k^* > 0 \Rightarrow \grad f(x^*)_k = \min_{1\leq i\leq n}\grad f(x^*)_i
\]

\clearpage
\section*{Problem 3}
Consider the function $f:\real^{n \times n}\times \real^n \rightarrow \real$ defined by $f(X, y) = y\T X^{-1}y$, where $X\in \real^{n\times n}$ and $y\in\real^n$.

(a) Suppose that $\dom f=\{(X, y) : X\in\symm^n_{++}\}$. Is $f(X, y)$ convex? If so, prove it. If not, give a counter-example.

(b) Suppose that $\dom f=\{(X, y): X+X\T \in\symm^n_{++}\}$. Is $f(X, y)$ convex? If so, prove it. If not, give a counter-example.



\clearpage
\section*{Problem 4}
Let $f:\real^n\rightarrow \real$ be a convex function with $f(0)\leq 0$.

(a) Show that the perspective function $tf(x/t)$ with domain $\{(x, t): t>0, x/t\in\dom f\}$ is nonincreasing as a function of $t$ for a fixed $x$.

(b) Let $g:\real^n\rightarrow \real$ be concave and positive on its domain. Show that the following function is convex:
\[
h(x) = g(x)f(x/g(x)), \quad \dom h=\{x\in\dom g: x/g(x) \in \dom f\}
\]

(c) Using the result of part (b), show that the following function is convex:
\[
h(x) = \frac{\|x\|_2^2}{(\prod_{i=1}^{n}x_i)^{1/n}},\quad \dom h = \real_{++}^n
\]

\clearpage
\section*{Problem 5}
(a) Let $X = \{x_1, x_2, \dots, x_m\}$ be a set of points in $\real^n$, where $m\geq n+2$. Show that $X$ can be partitioned in two sets $Y$ and $Z = X\backslash Y$ such that $\Co Y \cap \Co Z$ is non-empty. Here, $\Co Y$ and $\Co Z$ refer to the convex hulls of $Y$ and $Z$, respectively.

(b) Let $S_1, \dots, S_m$ be a collection of convex sets in $\real^n$, where $m\geq n + 2$. Suppose that the intersection of every $m-1$ sets from this collection is non-empty. Then the intersection of all sets $S_1\cap\dots\cap S_m$ is non-empty.



\end{document}
